---
title: "Stat641 Final Prep (HW#7 Redo)"
output: pdf_document
---
## 1
A researcher presents you with the following data Yi, i = 1, 2, . . . 29, which is independent and
identically normally distributed with a mean µ and variance σ^2

## 1.1  Set up the rejection region for testing the hypotheses H0 : µ = 20 vs H1 : µ 6= 20, based on Y1, . . . , Y29
and α = 0.05. Assume σ is unknown.


## 1.2 Calculate the power of your test for the following values of the true parameter µ:
19.9, 19.95, 19.99, 20, 20.05, 20.1, 20.15, 20.2, 20.25, 20.3, 20.4, 20.5.
The researcher provides the estimate, ˆσ = .43. This value of σ should just be used in the power
calculation and not in the actual testing procedure. Use your results to sketch a power curve for your
test. Be sure to label your axes clearly.

```{r}
mus = c(19.9, 19.95, 19.99, 20, 20.05, 20.1, 20.15, 20.2, 20.25, 20.3, 20.4, 20.5)

tprod_power_twosideed(mus, 20,0.43,n=29,alpha=0.05)

```

## 1.3 Determine the necessary sample size so that the power at µ = 20.15 is at least 0.80 for a level α = .05
test of H0 : µ ≤ 20 vs H1 : µ > 20.


```{r}
power.t.test(n=,delta = 20.15-20, sd=0.43,sig.level = 0.05,power = .80,type = c("one.sample"),alternative = "one.sided")

```



## 2
A new additive has been formulated to reduce the reaction time in a chemical process. With the
previously used additive, the average reaction time was 10 minutes. In order to evaluate the effectiveness of
the new additive, 15 batches of the material are formulated and the new additive is placed in the batches.
From previous studies, reaction times appear to have a normal distribution.

### 2.1
The mean reaction time from the 15 batches was 8.7 minutes with a standard deviation of 2 minutes.
Is there significant evidence using an α = .01 test that the average reaction time has been reduced?
Include the p-value with your decision.


```{r}

mo = 10
x_bar = 8.7
sigma_hat = 2
n = 15
alpha = 0.01

Tx = sqrt(n)*(x_bar-mo)/sigma_hat 

pt(Tx,n-1)

```
Based on the p-value, we would fail to reject the null.  Similarly, we could use the test statistic:

```{r}

qt(alpha,n-1)
Tx

Tx-qt(alpha,n-1)

c(x_bar,mo-qt(1-alpha,n-1)*sigma_hat/sqrt(n))

```


2.2 The process engineer had claimed that the new additive will reduce the average reaction time by at
least 1.5 minutes. What is the probability that the experiment will be able to detect a reduction of
the average reaction to 8.5 minutes or smaller using α = .01?


```{r}
power.t.test(n = 15, delta = 1.5, sd = 2, sig.level = 0.01, power =, alternative = c("one.sided"), type = c("one.sample"))

```

alternative method to calculate power
```{r}

pt(qt(0.01,14),14,sqrt(15)*(8.5-10)/2)

```



2.3  A new study is to be designed. What sample size is needed for an α = .05 test to have at least an 80%
chance to detect that the average reaction time is 9 minutes or less?

```{r}

power.t.test(n = , delta = 1, sd = 2, sig.level = 0.05, power = 0.80, type = c("one.sample"), alternative = c("one.sided"))

```



3.) A new device has been developed which allows patients to evaluate their blood sugar levels.
The most widely device currently on the market yields widely variable results. The new device is evaluated
by 25 patients having nearly the same distribution of blood sugar levels yielding the following data:

```{r}
data = c(125, 123, 117, 123, 115, 112, 128, 118, 124, 111, 116, 109, 125, 120, 113, 123, 112, 118,
121, 118, 122, 115, 105, 118, 131)
 

```


3.1 Is there significant evidence (α = .10) that the standard deviation in the readings from the new device
is less than 10?

```{r}

# first check if the data is normally distributed

shapiro.test(sort(data))

# p-value
TS = (length(data)-1)*var(data)/10^2

pchisq(TS,length(data)-1)


```

3.2 Compute the probability of a Type II error in using your test from part 1. for the following values of
σ: 5, 6, 7, 8, 9, 10

```{r}
n = length(data)
alpha = 0.10
sigma0 = 10
sigma = seq(5,10)
power = pchisq(((sigma0^2)/(sigma^2))*qchisq(alpha,n-1),n-1)
TypeIIerror = 1-power
cbind(sigma,TypeIIerror)

```


4.) (10 points) Refer to the blood sugar device data in Problem 3.
4.1 Is there significant (α = .05) evidence that median blood sugar readings was less than 120 in the
population from which the 25 patients were selected? Use the sign test and report the p-value.

```{r}

med0=120
x = data-med0
n = length(x)-sum(data==med0)

s_plus = sum(x>0)

data[which(x>0)]

pbinom(s_plus,size = n,prob = .5)

```

4.2 Is there significant (α = .05) evidence that median blood sugar readings was less than 120 in the
population from which the 25 patients were selected? Use the Wilcoxon signed rank test and report
the p-value.


```{r}

data = c(125, 123, 117, 123, 115, 112, 128, 118, 124, 111, 116, 109, 125, 120, 113, 123, 112, 118,
121, 118, 122, 115, 105, 118, 131)
med0 = 120



wilcox.signed.rank.test.one = function(y,med0,alpha,alternative = c("g","l","ne")){
  
  x = y-med0
  x = x[-which(x==0)]
  n = length(x)
  Group = c(x>0)

  grouped = as.data.frame(cbind(x,Group))
  grouped=grouped[order(abs(grouped[,1])),]
  grouped$rank = seq(1:length(grouped[,1]))

  for(i in abs(grouped$x)){
    rank_temp = grouped$rank[which(abs(grouped$x) == i)]
    if(length(rank_temp>1)){
      grouped$rank[which(abs(grouped$x) == i)] = sum(rank_temp)/length(rank_temp)
    }
  }
  wp = sum(grouped$rank[which(grouped$Group == 1)])
  wn = sum(grouped$rank[which(grouped$Group == 0)])
  wm = max(wp,wn)
  
  if(alternative == c("g")){
    critval = qsignrank(alpha,n,FALSE)
    pval = psignrank(wp-1,n,FALSE)
    true_level = psignrank(qsignrank(alpha,n,FALSE)-1,n,FALSE)
    
    
    return(cat("Reject Null if W+ =",wp,">",critval, 
               "\nP-value = ",pval,
               "\nTrue level of Test=",true_level))
    
  }else if(alternative == c("l")){
    critval = qsignrank(alpha,n,TRUE)
    pval = psignrank(wp,n,TRUE)
    true_level = psignrank(qsignrank(alpha,n,TRUE),n,TRUE)
    
    return(cat("Reject Null if W+ =",wp,"<",critval, 
               "\nP-value = ",pval,
               "\nTrue level of Test=",true_level))
    
  }else if(alternative == c("ne")){
    critval = qsignrank(alpha/2,n,FALSE)
    pval = 2*psignrank(wm-1,n,FALSE)
    true_level = 2*psignrank(qsignrank(alpha/2,n,FALSE)-1,n,FALSE)
    
    return(cat("Reject Null if Wmax =",wm,">",critval, 
               "\nP-value = ",pval,
               "\nTrue level of Test=",true_level))
    
  }
  
  
  
}



wilcox.signed.rank.test.one(y = data,med0, alpha = 0.05, alternative = c("l"))


x = x[-c(which(x==0))]
y = rep(med0, times = length(x))

wilcox.test(x,y, alternative = "l",paired = T)

```



4.3 Place a 90% upper bound on the median blood sugar reading.





5.) The current method of identifying patients at risk of sudden cardiac death can be identified
with 80% accuracy. A change in the method has hopefully improved the accuracy. To evaluate the new
method, 100 people are tested and the new method produced the result on 92 of the 100 people.


5.1 Place a 95% confidence interval on the accuracy of the device.

```{r}

### Agrestic Coull CI

agresti.coullCI = function(y,n,alpha){

  if((n < 40) | min(n*(y/n),n*(1-(y/n))) < 5){
    
    return("Assumptions are not met, Agresti-Coull CI is not appropriate")
  }else{
    y_tild = y+.5*(qnorm(1-alpha/2)^2)
    n_tild = n+qnorm(1-alpha/2)^2

    p_tild = y_tild/n_tild

    CI = c(p_tild - qnorm(1-alpha/2)*((sqrt(p_tild*(1-p_tild)))/sqrt(n)),p_tild +  qnorm(1-alpha/2)*((sqrt(p_tild*(1-p_tild)))/sqrt(n))) 
 
return(CI)
    
    
  }
  
}


agresti.coullCI(y = 92, n = 100, alpha = .05)

```


5.2 Is there substantial evidence (α = .05) that the improved method has increased the accuracy over the
current method?

```{r}

test.pop.proportion = function(y="",p_hat = "",n,p0,alpha, alternative = c("l","g","ne")){
  if(p_hat == ""){
    p_hat = y/n
  }
  
  if(y == ""){
    y = p_hat*n
  }
  
  if((min(p_hat*n,n*(1-p_hat)) < 5) | n<40){
    
    if(alternative == c("g")){
      
      critval = qbinom(1-alpha,n,p0)
      p_val = 1-pbinom(y-1,n,p0)
      
      return(cat("Reject Null if Y=",y,">",critval,
                 "\nP-value =",p_val))
      
    }else if(alternative == c("l")){
      
      critval = qbinom(alpha,n,p0)
      p_val = pbinom(Y,n,p0) 
      
      return(cat("Reject Null if Y=",y,"<",critval,
                 "\nP-value =",p_val))
      
    }else if(alternative == c("ne")){
      
      critval = c(qbinom(alpha/2,n,p0),qbinom(1-alpha/2,n,p0))
      p_val = 2*min(pbinom(y,n,p0),1-pbinom(y-1,n,p0))
      
      return(cat("Reject Null if Y=",y,"<",critval[1],"or Y =",y,">",critval[2],
                 "\nP-value =",p_val))
      
    }
    
    
    
  }else{
    
    TS = (p_hat - p0)/sqrt((p0*(1-p0))/n)
    
    if(alternative == c("g")){
      critval = qnorm(1-alpha)
      p_val = 1-pnorm(TS)
      
      return(cat("Reject Null if TS=",TS,">",critval,
                 "\nP-value =",p_val))
      
    }else if(alternative == c("l")){
      
      critval = qnorm(alpha)
      p_val = pnorm(TS)
      
      return(cat("Reject Null if TS=",TS,"<",critval,
                 "\nP-value =",p_val))
      
    }else if(alternative == c("ne")){
      
      critval = qnorm(1-alpha/2)
      p_val = 2*(1-pnorm(abs(TS)))
      
      return(cat("Reject Null if |TS|=",abs(TS),">",critval,
                 "\nP-value =",p_val))
      
      
    }
    
    
  }
  
}


test.pop.proportion(y=92,n=100,p0=0.8,alpha = .05, alternative = c("g"))

```


5.3 Compute the power of the test in part 2. to detect that the accuracy of the improved method is 75%,
80%, 85%, 90%, 95%.


```{r}

ps = seq(.75,.95,by=.05)
ys = 1-pbinom(qbinom(.95,100,.8)-1,100,ps)
cbind(ps, ys)

qbinom(.95,100,.8)

```


5.4 How many patients would need to be included in a new study in order to have a power value of 80%
if the new method had an accuracy of 90%.

```{r}

### sample size calucaltion for population proportion p

sample.size.pop.prop = function(p0,p1,alpha,beta){
  
  delta = abs(p1-p0)
  z_a = abs(qnorm(alpha))
  z_b = abs(qnorm(beta))
  
  n = ceiling((((z_a*sqrt(p0*(1-p0)))+(z_b*sqrt(p1*(1-p1))))/delta)^2)
  
  return(n)
}


sample.size.pop.prop(p0 = .8,p1=.9, alpha = .05, beta = .80)


```

(MC5.) A researcher wants to determine if there is an increase in the likelihood that people will purchase a product
after a redesign of the product. The current market share is 20%. Initially, the researcher was planning on
using a random sample of n=20 persons with an α = .05 test to evaluate the product. He wants you to
calculate the chance that the study will fail to detect that preference for the product has been increased if
in fact the preference for the new product is 40%. This chance is

```{r}

pbinom(qbinom(1-.05,20,.20),20,.40)

```

(MC6)  A random sample of n=15 from a normally distributed population is used to construct a level α = .01 test
of Ho : µ ≤ 20 versus H1 : µ > 20, where µ is the mean of the population. The probability of a Type II
error for µ > 20 + .8σ is at most


```{r}
typeIIerrorprob = pt(qt(1-.01, 15 - 1), 15 - 1, sqrt(15)*.8)
typeIIerrorprob

```
(MC7) A psychologist is investigating the IQ level of young children who have been in a head start program. She
wants to determine if the variation in IQ scores for the population of head start students is smaller than
the variation in the general population of children under the age of 6 which has a variation of σ = 10.2.
She also informs you that the distribution of IQ scores is highly right skewed. Suppose she uses the test:
reject Ho is





