---
title: "STAT 608 - Assignment1"
author: "Jack Rodoni"
date: "1/28/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# 1. Question 1, page 38 in our textbook. Notice that to answer these questions correctly, you should be thinking like a statistician and talking about population parameters, not only sample statistics. That is, do some inference in every part, using as much everyday layperson’s terminology as possible. For example, part b should not just say “The intercept is (or is not) 10,000.” What does an intercept mean in context to someone who is selling movie tickets? Use the discussion board as needed to get all your details correct.

  1. The web site www.playbill.com provides weekly reports on the box office 
ticket sales for plays on Broadway in New York. We shall consider the data 
for the week October 11–17, 2004 (referred to below as the current week). 
The data are in the form of the gross box office results for the current week 
and the gross box office results for the previous week (i.e., October 3–10, 
2004). The data, plotted in Figure 2.6 , are available on the book web site in 
the file playbill.csv. 

  Fit the following model to the data: Y= b~0~ +b~1~X  where Y is the gross box 
office results for the current week (in $) and x is the gross box office results for the 
previous week (in $). Complete the following tasks:

```{r}

playbill <- read.csv(paste0("C:/Users/jackr/OneDrive/Desktop/Graduate School ",
                     "Courses/STAT 608 - Regression Analysis/Data/playbill.csv"))



lm1 = with(playbill,lm(CurrentWeek~LastWeek))

modelsum = summary(lm1)
summary(lm1)
equatiomatic::extract_eq(lm1, use_coefs = TRUE)

```


  (a) Find a 95% confidence interval for the slope of the regression model, b~1~.  Is 1 a plausible value for b~1~? Give a reason to support your answer.

```{r eval=FALSE}
# calculation done by R
confint(lm1, level = 0.95)

# if we just want B1
confint(lm1, "playbill$LastWeek", level = 0.95)

```

```{r eval=FALSE}
# By Hand calculation
c(lm1$coefficients[2] - qt(.975,16)*modelsum$coefficients[2,2],
  lm1$coefficients[2] + qt(.975,16)*modelsum$coefficients[2,2])

# or
SXX = sum((playbill$LastWeek-mean(playbill$LastWeek))^2)

lm1$coefficients[2] +c(-qt(.975,16)*(modelsum$sigma/sqrt(SXX)),qt(.975,16)*(modelsum$sigma/sqrt(SXX)))


```

```{r echo=FALSE}
confint(lm1, level = 0.95)

```


**Yes, 1 is a reasonable value for B~1~ as 1 is contained in our confidence interval**

  (b) Test the null hypothesis H~0~ : b~0~ = 10000 against a two-sided alternative. 
Interpret your result. 


```{r}
teststat = (lm1$coefficients[1] - 10000)/modelsum$coefficients[1,2]
# p-value calculation
2*(1-(pt(abs(teststat),16))) # See STAT 641 H.O 12 pg 24

# or just compare test stat to critical value
crit.val = qt(0.975,16)

abs(teststat)>crit.val


```

**The p-value is 0.7517807 -> we would fail to reject the null that b~0~ = 10000**


  (c) Use the fitted regression model to estimate the gross box office results for 
the current week (in \$) for a production with \$400,000 in gross box office 
the previous week. Find a 95% prediction interval for the gross box office. 
is \$450,000 a feasible value for the gross box office results in the current week, for a production with  \$400,000 in gross box office the previous week? Give a reason to support your answer.



```{r results="hide"}
# 95% CI
summary(lm1)
## R calculation
predict(lm1, newdata = data.frame(LastWeek = 400000), interval = "confidence")
## by hand
with(playbill, 399637 + c(-1, 1) * qt(0.975, 16) * 18010 * 
  sqrt((1/18) + (400000 - mean(LastWeek)) ^ 2 / sum((LastWeek - mean(LastWeek)) ^ 2)))


# 95% PI
## R calculation
predict(lm1, newdata = data.frame(LastWeek = 400000), interval = "prediction")
## by hand
PI = with(playbill, 399637 + c(-1, 1) * qt(0.975, 16) * 18010 * 
  sqrt(1+(1/18) + (400000 - mean(LastWeek)) ^ 2 / sum((LastWeek - mean(LastWeek)) ^ 2)))
PI

```

**450,000 is not a feasible value for the gross box office results in the current week  for a production with $400,000 in gross box office the previous week**


```{r echo=FALSE}
PI
```


(d) Some promoters of Broadway plays use the prediction rule that next week’s 
gross box office results will be equal to this week’s gross box office results. 
Comment on the appropriateness of this rule.

**This rule seems reasonable as our last weeks value is included in our 95% prediction interval for next week**

# 2. Show that Var(Y~i~|X~i~ = x~i~) = Var($\epsilon~i~) in the simple linear regression model. Don’t overthink this; the answer is simple. What did you assume in answering this?


\newpage

# 3. Define using only words what the least squares criterion is.

Simply put, the least squares criterion is when we are fitting a regression model we select the model that minimizes the sum of the square errors.  In other words, we minimize the sum of the squares of the projections from our data to the model space.  


          

# 4. Question 4, page 40 in our textbook, except do:

  4. Straight-line regression through the origin:
  In this question we shall make the following assumptions:
  (1) Y is related to x by the simple linear regression model:
      Y~i~ = BX~i~+e~i~ (i = 1,2,3,...)
      i.e. E(Y|X=x~i~) = BX~i~
  (2) The errors e~1~,e~2~,...,e~n~ are independent of each other
  (3) The errors e~1~,e~2~,...,e~n~ have a common variance  sigma^2^ 
  (4) The errors are normally distributed with a mean of 0 and variance  sigma^2^ (especially when the sample size is       small), i.e., (e|X)\~\N(0, sigma^2^)         
  

  (a) Setup:
  i. Write down your design matrix X.
      \
      \
      \
      \
      \
      \
  ii. Show, using matrix notation and starting with the principle of least squares, that
      the least squares estimate of B is given by:
      \
      \
      \
      \
      \
      \
      \
      \
      \
      \
      \
      \
      \
      \
      \
      \
      \
      \
      \
      \
      \
      \
      \
      \
      \
    
  (b) Under the above assumptions show that:
  
  i.
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  
  ii.
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  \
  

      


      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      
      

          
          


















          
          
          
          
          
          
















































